Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2021-09-22 analysis/paper_draft2.Rmd
minor edits

And as our prototype app shows, if the historical strike data is collected and loaded into the app in advance, this approach can be easily used by attorneys and others in real time during jury selection. Despite _Batson_ and similar law, racial bias persists in the use of peremptory challenges [@Diamond2018], perhaps in part because attorneys fear that _raising_ a _Batson_ challenge raises with it the embarrassment and discomfort associated with accusing another attorney of being racist [@OBrien2019, 29; @Offit2021]. Whether an app like the prototype presented here might reduce that and other barriers to bringing _Batson_ challenges is a question worth studying further.

At the same time, the Bayesian approach here also requires more from those who would collect the appropriate historical strike data, given the choice of how to model a court's particular peremptory-challenge process. For example, given the model we used, we required data on the cognitive-class composition of the prospective jurors not just at the time of the first strike, but at each time either party used a strike. Moreover, suppose an attorney of interest exercised strikes in two courts, each with different strike procedures. In some cases, one court's strike procedure can be modeled as a special case of a more general model that applied to other court's procedure (as we showed above). If not, that attorney's historical strike data in one court could not used to estimate the posterior of that attorney's bias parameter, given a current trial in the other court.

Finally, for the approach here, like any bias-detection method relying on historical strike data, the ideal data, therefore, is, for each attorney, not just who they struck in the current trial, but a complete history of who they struck in all their past cases. The problem: Jury-selection records in past cases are hard to get because of court practices and otherwise require painstaking effort to compile [@Grosso2017; @Wright2018]. Possible causes include that some jurisdictions do not require recording relevant or detailed enough information about jurors who were or could have been struck (e.g., their race or gender). This forces researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information. Finally, judges and court staff vary in how often they fully complete or maintain the information they are required to record. This presents an issue for inferring bias if historical strike data is missing not at random. These data challenges, however, might decrease over time, as judges, attorneys, and jury clerks realize that more complete historical strike data makes it easier, with the approach described here, to credibly infer or disprove bias under _Batson_ and similar law.

>>>>

And as our prototype app shows, if the historical strike data loaded into the app in advance, attorneys and others can the approach here in real time during jury selection. Despite _Batson_ and similar law, racial bias persists in the use of peremptory challenges [@Diamond2018], perhaps in part because attorneys fear that _raising_ a _Batson_ challenge raises with it the embarrassment and discomfort associated with accusing another attorney of being racist [@OBrien2019, 29; @Offit2021]. It is worth studying whether an app like the prototype presented here might reduce that and other barriers to bringing _Batson_ challenges.

At the same time, the Bayesian approach here also requires more from those who would collect the appropriate historical strike data. For example, given the model we used, we required data on the cognitive-class composition of the prospective jurors not just at the time of the first strike, but at each time either party used a strike. Moreover, suppose an attorney of interest exercised strikes in two courts, each with different strike procedures. In some cases, one court's strike procedure can be modeled as a special case of a more general model that applied to other court's procedure (as we showed above). If not, that attorney's historical strike data in one court could not used to estimate the posterior of that attorney's bias parameter, given a current trial in the other court.

Finally, for the approach here, like any bias-detection method relying on historical strike data, the ideal data is, for each attorney, a complete history of who they struck in all their past cases. The problem: Jury-selection records in past cases are hard to get because of court practices and other barriers [@Grosso2017; @Wright2018]. For example, some jurisdictions do not require recording relevant or detailed enough information about prospective jurors who were or could have been struck (e.g., their race or gender). This forces researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information. And judges and court staff vary in how often they fully complete or maintain the information they are required to record. Incomplete historical strike data resents an issue for inferring bias if such data is incomplete or missing not at random. These data challenges, however, might decrease over time as judges, attorneys, and jury clerks realize that more historical strike data makes it easier, with the approach described here, to credibly infer or disprove bias under _Batson_ and similar law.
